Improving the Replicability of Psychological Science Through Pedagogy

### 2016-09-23 paper/paper.Rmd
spell check

We chose to use Amazon Mechanical Turk (AMT) as a platform for our replication studies. This choice was made to facilitate the process of recruiting samples large enough to enable replications of between-subjects designs (which typically require large samples) as well as to help students learn about a valuable resource for recruitment whose use requires some specialized knwoledge. Our general class structure does not depend on the use of AMT, however, and the use of this particular online sample should be viewed as limiting the interpretation of some results.

@@

Repeat administration of experiments on AMT has been raised as an issue in some prior work [e.g., @chandler2014], but tracking participation in specific paradigms is an open challenge. We did not ask participants whether they had participated in similar research previously, however. First, half of the AMT population is estimated to change every six months [@stewart2015], and we expected that most of the studies in our sample that had been conducted on AMT had been performed at least a year previously (though see SOM for discussion of one particular paradigm that has been used more recently). Second, we suspected that asking this sort of question would lead to a large number of innaccurate responses.

@@

While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent concerns that using confidence intervals in this manner may lead to misleading interpretations [@morey2016], as well as discrepencies in the authors' reporting of effect sizes across the articles we included (which made computing confidence intervals difficult in some cases).

@@

One of the barriers to more reproducible work in psychology is pedagogical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web experiments or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environemnt -- students can quickly learn to use a reproducible toolset.

@@

These statistical findings mask a more optimistic message, however: In most of our projects, the next step for a motivated experimenter is clear. For example, in some projects we suspect that a followup could find strong evidence for the phenomenon of interest by titrating the particular planned analysis or the difficulty of the stimulus materials [@liverence2015;@storm2015;@xu2015]. In others, the difficulty appeared to be statistical power, since differences were in the predicted direction and sometimes reached significance in subsidiary analyses, so a followup would likely require a larger sample [@zaval015;@lewis2015;@scopelliti2015]. And in some, differences of population would mean that followups would require further stimulus or task adaptation [@sofer2015;@wang2015]. More generally, we believe that our work here underscores the importance of iterated replication [@lewis2016].

Many statistical factors leading to lowered replicability have been discussed in past work, including analytic flexibility, context dependency, publication bias, and low statistical power among others [e.g., @button2013; @vanbavel2016; @ioannidis2005]. For both reasons of design and scale, our study here cannot disentangle these factors empirically. Nevertheless, the decreases in effect size and evidential value we report seem consistent with two explanations. First, we likely saw a "winner's curse," with initial publications tending to over-estimate effects. Second, we also see evidnece that specific details of studies vary from context to context; creating successful adaptations can be challenging [@vanbavel2016]. Both of these explain our observations, and we suspect both are present to some degree.

@@

This paper investigated motivations to save for retirement, and specifically whether changing participants' relationships to the future (making it feel more imminent) would lead them to report that they would begin saving for retirement sooner. Our target was Study 4, which found that seeing the time to retirement in days (10,950) rather than years (30) caused participants to say they would begin saving sooner, even controlling for age, education, and income in a linear model (there was also a manipulation of whether saving was incremental that did not result in an effect). We failed to find a significant relationship between the time metric manipulation and saving time when including demographic controls (the key test we selected). However, when we did not include demographic controls, we did find a marginally-significant relationship. In exploratory analyses, we also observed an un-predicted effect of income such that participants who reported higher income selected to start saving sooner.

>>>>

We chose to use Amazon Mechanical Turk (AMT) as a platform for our replication studies. This choice was made to facilitate the process of recruiting samples large enough to enable replications of between-subjects designs (which typically require large samples) as well as to help students learn about a valuable resource for recruitment whose use requires some specialized knowledge. Our general class structure does not depend on the use of AMT, however, and the use of this particular online sample should be viewed as limiting the interpretation of some results.

@@

Repeat administration of experiments on AMT has been raised as an issue in some prior work [e.g., @chandler2014], but tracking participation in specific paradigms is an open challenge. We did not ask participants whether they had participated in similar research previously, however. First, half of the AMT population is estimated to change every six months [@stewart2015], and we expected that most of the studies in our sample that had been conducted on AMT had been performed at least a year previously (though see SOM for discussion of one particular paradigm that has been used more recently). Second, we suspected that asking this sort of question would lead to a large number of inaccurate responses.

@@

While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent concerns that using confidence intervals in this manner may lead to misleading interpretations [@morey2016], as well as discrepancies in the authors' reporting of effect sizes across the articles we included (which made computing confidence intervals difficult in some cases).

@@

One of the barriers to more reproducible work in psychology is pedagogical: Some of the tools (programmatic statistical analysis, version control) can feel inaccessible or intimidating to beginners. Indeed, our students, who came from a variety of disciplinary backgrounds, had relatively little experience with web experiments or literate, reproducible data analysis. Yet by the end of the course, all had gained sufficient proficiency with the suite of technical and conceptual tools necessary to carry out a study independently. This experience suggests that it is possible to convey key concepts in sufficient depth that -- within a supportive environment -- students can quickly learn to use a reproducible toolset.

@@

These statistical findings mask a more optimistic message, however: In most of our projects, the next step for a motivated experimenter is clear. For example, in some projects we suspect that a followup could find strong evidence for the phenomenon of interest by titrating the particular planned analysis or the difficulty of the stimulus materials [@liverence2015;@storm2015;@xu2015]. In others, the difficulty appeared to be statistical power, since differences were in the predicted direction and sometimes reached significance in subsidiary analyses, so a followup would likely require a larger sample [@zaval2015;@lewis2015;@scopelliti2015]. And in some, differences of population would mean that followups would require further stimulus or task adaptation [@sofer2015;@wang2015]. More generally, we believe that our work here underscores the importance of iterated replication [@lewis2016].

Many statistical factors leading to lowered replicability have been discussed in past work, including analytic flexibility, context dependency, publication bias, and low statistical power among others [e.g., @button2013; @vanbavel2016; @ioannidis2005]. For both reasons of design and scale, our study here cannot disentangle these factors empirically. Nevertheless, the decreases in effect size and evidential value we report seem consistent with two explanations. First, we likely saw a "winner's curse," with initial publications tending to over-estimate effects. Second, we also see evidence that specific details of studies vary from context to context; creating successful adaptations can be challenging [@vanbavel2016]. Both of these explain our observations, and we suspect both are present to some degree.

@@

This paper investigated motivations to save for retirement, and specifically whether changing participants' relationships to the future (making it feel more imminent) would lead them to report that they would begin saving for retirement sooner. Our target was Study 4, which found that seeing the time to retirement in days (10,950) rather than years (30) caused participants to say they would begin saving sooner, even controlling for age, education, and income in a linear model (there was also a manipulation of whether saving was incremental that did not result in an effect). We failed to find a significant relationship between the time metric manipulation and saving time when including demographic controls (the key test we selected). However, when we did not include demographic controls, we did find a marginally-significant relationship. In exploratory analyses, we also observed an unpredicted effect of income such that participants who reported higher income selected to start saving sooner.
